\input{./._relpath-to-latexroot.ltx}
\documentclass[\latexroot/\projectname]{subfiles}
\input{\latexroot/@resources/subfile-setup.ltx}
\appendix\setcounter{section}{0} % standalone: use appendix numbering and start at B (whenstandalone removed - redundant in preamble)

\begin{document}

\FloatBarrier
% This will be 'A Details of the HANK and SAM Model' as the first appendix section

% Always use starred section to suppress automatic bookmark
% The bookmark will be created in HAFiscal.tex at the stub page location
\section{\href{https://econ-ark.github.io/HAFiscal/\#sec:appendix-hank}{Details of the HANK and SAM Model}}
\whenintegrated{\label{sec:appendix-hank}} % integrated doc: SST for labels

% Old BuildWithoutNoSplurge logic removed - OFFLINE appendices are always included in full

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:hank-households}{Households}}\whenintegrated{\label{sec:hank-households}}

The household block follows closely to the main text with a few exceptions. First, the splurge only occurs out of equilibrium---that is, the steady state of the model is calculated without the splurge behavior. Second, the level of permanent income of all newborns is equal to one. Furthermore, all households face the same employment to unemployment and unemployment to employment probabilities. The probabilities are calibrated to the transition probabilities of high school graduates from the main text. Lastly, following the notation of~\cite{arsJumpsHumps}, $r^{a}_{t}$ will denote the economy wide ex-ante real interest rate.

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:goods-market}{Goods Market}}\whenintegrated{\label{sec:goods-market}}

A continuum of monopolistically competitive intermediate goods producers, indexed by \( j \in [0,1] \), produces intermediate goods \( Y_{jt} \), which are sold to a final goods producer at price \( P_{jt} \). Each period, these producers fully consume their profits.

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:hank-final-goods-producer}{Final Goods Producer}}\whenintegrated{\label{sec:hank-final-goods-producer}}

A perfectly competitive final goods producer purchases intermediate goods \( Y_{jt} \) from intermediate good producer $j$ at price \( P_{jt} \) and produces the final good \( Y_t \) using a CES production function:

\[
  Y_{t} = {\left(\int_{0}^{1} Y_{jt}^{\frac{\epsilon_{p}-1}{\epsilon_{p}}}\, dj\right)}^{\frac{\epsilon_{p}}{\epsilon_{p}-1}},
\]
where $\epsilon_{p}$ is the elasticity of substitution.

Given $P_{jt}$, the price of intermediate good $j$, the final goods producer maximizes profits by solving:
\[
  \max_{Y_{jt}} P_{t} {\left(\int_{0}^{1} Y_{jt}^{\frac{\epsilon_{p}-1}{\epsilon_{p}}}\, dj\right)}^{\frac{\epsilon_{p}}{\epsilon_{p}-1}} - \int_{0}^{1} P_{jt} Y_{jt} ,\ dj.
\]
%\vspace{.2cm}

The first order condition leads to demand for good $j$ given by
\[
  Y_{jt} = {\left(\frac {P_{jt}}{P_{t}}\right)}^{- \epsilon_{p}} Y_{t},
\]
and the price index
\[
  P_{t} = {\left(\int_{0}^{1} P_{jt}^{1-\epsilon_{p}}\,dj \right )}^{\frac{1}{1-\epsilon_{p}}}.
\]

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:hank-intermediate-goods-producers}{Intermediate Goods Producers}}\whenintegrated{\label{sec:hank-intermediate-goods-producers}}

Intermediate goods producers produce according to a production function linear in labor~$L_{t}$:
\[
  Y_{jt} =  Z L_{jt},
\]
where $Z$ is total factor productivity.
%\vspace{.3cm}

Each intermediate goods producer hires labor \( L_t \) from a labor agency at cost \( h_t \). Given this labor cost, each producer sets \( P_{jt} \) to maximize profits while facing price stickiness à la~\cite{Rotemberg1982}. In HANK models with sticky prices, profits tend to be countercyclical, and when households have high MPCs, this can generate countercyclical consumption responses to dividends. To simplify, we assume that intermediate goods producers fully consume their profits rather than distributing them to households, thereby abstracting from consumption responses to firm profits. Each producer maximizes profits by solving:

\[
  J_{t}\left(P_{jt}\right) = \max_{\{P_{jt}\}} \left\{\frac{P_{jt}Y_{jt}}{P_{t}} - h_{t} L_{jt} -  \frac{\varphi}{2}{\left( \frac{P_{jt} - P_{jt-1}}{P_{jt-1}} \right)}^{2} Y_{t}  + J_{t+1}\left(P_{jt+1}\right) \right\},
\]

where $\varphi$ determines the cost of adjusting the price and, hence, the degree of price stickiness.

The problem can be rewritten as the standard New Keynesian maximization problem:
\[
  \max_{\{P_{jt}\}} \mathrm{E}_{t}\left[\sum_{s=0}^{\infty}  M_{t,t+s} \left( \left( \frac{P_{jt+s}}{P_{t+s}} - MC_{t+s}\right)Y_{jt+s} -  \frac{\varphi}{2}{\left( \frac{P_{jt+s}}{P_{jt+s-1}} - 1\right)}^{2} Y_{t+s} \right)\right],
\]
where $MC_{t} = \frac{h_{t}}{Z}$.

Given that all firms face the same adjustment costs, there exists a symmetric equilibrium where all firms choose the same price with $P_{jt}=P_{t}$ and $Y_{jt}=Y_{t}$.

The resulting Phillips Curve is
\[
  \epsilon_{p} MC_{t} = \epsilon_{p} - 1 + \varphi ( \Pi_{t} -1) \Pi_{t} - M_{t,t+1} \varphi (\Pi_{t+1} -1 ) \Pi_{t+1} \frac{Y_{t+1}}{Y_{t}}
\]
where $\Pi_{t} = \frac{P_{t}}{P_{t+1}}$.

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:hank-labor-market}{Labor market}}\whenintegrated{\label{sec:hank-labor-market}}

A risk-neutral labor agency supplies labor \( N_t \) to intermediate goods producers at cost \( h_t \) by hiring households at wage \( w_t \). To hire workers, the agency posts vacancies \( v_t \), which are filled with probability \( \phi_t \). Household job search is random. Following~\cite{Bardoczy2022}, we assume the labor agency cannot observe individual household productivity. Instead, it only observes the average productivity of all employed workers, which is normalized to one.

\textbf{Labor agency.} The labor agency determines how many vacancies to post and how much labor to sell by solving the following problem:
\[
  J_{t}(N_{t-1})  = \max_{N_{t},v_{t}} \left\{( h_{t} - w_{t}) N_{t}- \kappa v_{t} + \mathrm{E_{t}}\left[ \frac{J_{t+1}(N_{t})}{1 + r^{a}_{t}}\right]\right\},
\]
subject to
\[
  N_{t} = (1-\omega)N_{t-1} + \phi_{t} v_{t}.
\]

The parameters $\kappa$ and $\omega$ are, respectively, the cost of posting a vacancy, and the job separation rate.

The resulting job creation curve is:
\[
  \frac{\kappa}{\phi_{t}}  = (h_{t} - w_{t})+  (1-\omega)\mathrm{E_{t}}\left[   \frac{\kappa}{(1+r^{a}_{t}) \phi_{t+1}} \right].
\]

\textbf{Matching.} The matching process between households and the labor agency follows a Cobb-Douglas matching function:
\[
  m_{t} = \chi e_{t}^{\alpha} v_{t}^{1-\alpha},
\]
where $m_{t}$ is the mass of matches, $e_{t}$ is the mass of job searchers, $\alpha$ is the matching function elasticity, and $\chi$ is a matching efficiency parameter.

The vacancy filling probability \( \phi_t \) and the job finding probability \( \eta_t \) evolve as follows:
\[
  \eta_{t} = \chi \Theta_{it}^{1-\alpha}
\]
%$$\eta_{t}(X) = \chi q(X) \Theta_{it}^{1-\alpha} $$
\[
  \phi_{t} = \chi \Theta_{t}^{-\alpha}
\]
where $\Theta_{t} = \frac{v_{t}}{e_{t}}$ is labor market tightness.

\textbf{Wage Determination.} Following~\cite{Gornemann2021} and~\cite{Blanchard2010}, we assume the real wage evolves according to the following rule:
\[
  \log\left(\frac{w_{t}}{w_{ss}}\right)  = \phi_w \log\left( \frac{ w_{t-1}}{ w_{ss}} \right) +   (1 - \phi_w) \log\left( \frac{N_{t}}{N_{ss}}\right),
\]
where $\phi_w$ dictates the extent of real wage rigidity.

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:hank-fiscal-policy}{Fiscal Policy}}\whenintegrated{\label{sec:hank-fiscal-policy}}

The government issues long term bonds $B_{t}$ at price $q^{b}_{t}$ in period $t$ that pays $\delta^{s}$ in period $t+s+1$ for $s \in \{0,1,2,\ldots\}$.

The bond price satisfies the no arbitrage condition:
\[
  q^{b}_{t} = \frac{ 1  + \delta \mathrm{E}_{t}[q^{b}_{t+1}]}{1+r^{a}_{t}}.
\]

The government funds its expenditures through debt and taxes, subject to the following budget constraint:
\[
  (1 + \delta q^{b}_{t})B_{t-1} + G_{t}  + S_{t} = \tau_{t} w_{t} N_{t}+ q^{b}_{t}B_{t},
\]
where $S_{t}$ are payments for unemployment insurance and other transfers.

For all stimulus policies, except tax cuts, we follow~\cite{arsJumpsHumps} and allow the tax rate to adjust in order to stabilize the debt-to-GDP ratio:
\[
  \tau_{t} - \tau_{ss} = \phi_{B} q^{b}_{ss} \frac{B_{t-1} - B_{ss} }{Y_{ss}}
\]
where $\phi_{B}$ governs the speed of adjustment.

For the tax cuts, we assume government expenditures adjust following:
\[
  G_{t} - G_{ss} = \phi_{G} q^{b}_{ss} \frac{B_{t-1} - B_{ss} }{Y_{ss}}
\]
where $\phi_{G}$ governs the speed of adjustment of government spending in response to debt.

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:hank-monetary-policy}{Monetary Policy}}\whenintegrated{\label{sec:hank-monetary-policy}}

The central bank follows a standard Taylor rule that responds solely to inflation:
\[
  i_{t} = r^{*} +\phi_{\pi} \pi_{t},
\]
where $\phi_{\pi}$ is the coefficient on inflation. Inflation is given by $\pi_t = P_t/P_{t-1}-1$, and~$r^{*}$ is the steady state interest rate.

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:hank-equilibrium}{Equilibrium}}\whenintegrated{\label{sec:hank-equilibrium}}

An equilibrium in this economy is a sequence of:
\begin{itemize}[label=--]
  \item
        Policy Functions ${\left( c_{it}(m) \right )}_{t=0}^{\infty}$ normalized by permanent income.
  \item
        Prices $ {\left( r^{a}_{t+1}, i_{t}, q^{b}_{t},  w_{t}, h_{t} , \pi_{t} , \tau_{t} \right) }_{t=0}^{\infty}$.
  \item
        Aggregates $ {\left(C_{t}, Y_{t} , N_{t},   \Theta_{t},  B_{t} , A_{t}  \right)}_{t=0}^{\infty}$.
\end{itemize}

Such that:
\begin{itemize}[label=--]
  \item
        ${\left(c_{it}(m)\right)}_{t=0}^{\infty}$  solves the household's maximization problem given ${\left( w_{t}, \eta_{t},  r^{a}_{t} , \tau_{t} \right)}_{t=0}^{\infty}$.
  \item
        The final goods producer and intermediate goods producers both maximize their respective objective functions.
  \item
        The nominal interest rate is determined by the central bank's Taylor rule.
  \item
        The tax rate is set according to the fiscal rule, ensuring that the government budget constraint is satisfied.
  \item
        The value of assets is equal to the value of government bonds:
        \[
          A_t =  q^{b}_{t}B_{t}.
        \]
  \item
        The goods market clears\footnote{Note if profits were not held by firms then the goods market condition would be $ C_{t}  + G_{t}  = Y_{t} -  \kappa v_{t} - \frac{\varphi}{2}{\left( \frac{P_{t}}{P_{t-1}} - 1\right)}^{2} Y_{t}  $.  In particular, since firm profits are $D_{t} = Y_{t} -  w_{t} N_{t}  - \kappa v_{t} - \frac{\varphi}{2}{\left( \frac{P_{t}}{P_{t-1}} - 1\right)}^{2} Y_{t} $, then the goods market condition would become $ C_{t}  + G_{t}  =w_{t} N_{t}  + D_{t} = Y_{t} -  \kappa v_{t} - \frac{\varphi}{2}{\left( \frac{P_{t}}{P_{t-1}} - 1\right)}^{2} Y_{t}  $. }:
        \[
          C_{t}  = w_{t} N_{t}  + G_{t},
        \]
        where $C_{t} \equiv  \int_{0}^{1} \textbf{c}_{it} \, di$.
  \item
        The labor demand of intermediate goods producers equals labor supply of labor agency:
        \[
          L_{t} =  N_{t}.
        \]
\end{itemize}

\begin{table}
  \begin{center}
    \renewcommand{\arraystretch}{1.6}
    \caption{Calibration}\whenintegrated{\label{table:Calibration}} % integrated doc: SST for labels
    \makebox[\textwidth]{
      \fetchgeneratedtabular{\latexroot/tabular/Calibration.ltx}
    }
  \end{center}
\end{table}

\subsection{\href{https://econ-ark.github.io/HAFiscal/\#sec:hank-calibration}{Calibration of Non-Household Blocks }}\whenintegrated{\label{sec:hank-calibration}}

The elasticity of substitution is set to 6, and the price adjustment cost parameter is set to 96.9 as in~\cite{Ravn2017,Ravn2021}.
The vacancy cost is set to 7 percent of the real wage as in~\cite{Christiano2016}.\footnote{The range of plausible values lie between $4 percent$ and $14 percent$~\cite{Silva2009}} The matching elasticity is 0.65 following~\cite{Ravn2017,Ravn2021}. The job separation rate is set to 0.092. As in section~\ref{sec:calib}, we set the job finding probability in the steady state for the unemployed $\eta_{ss}$ to 0.67. Along with the job separation rate, this gives a probability of transitioning from employment to unemployment within a quarter of $3.1$ percent which is the value we use for the Highschool group in section~\ref{sec:calib}. The quarterly vacancy filling rate is 0.71 as in~\cite{DenHaan2000} (and together with our other choices, this pins down the matching efficiency $\chi$). The degree of wage rigidity $\phi_w$ is set to 0.837 following~\cite{Gornemann2021}. The tax rate is set to 0.3 and government spending is set to clear the government budget constraint. The parameters that dictate the speed of fiscal adjustment, $\phi_{B}$ and $\phi_{G}$, are set to 0.015, the lower bound of the estimates in~\cite{arsJumpsHumps}.\footnote{The speed of adjustment parameter is set to the lower bound to ensure that the policies evaluated in the HANK and SAM model are almost entirely deficit financed.} Furthermore, the decay rate of government coupons is set to $\delta = 0.95$ to match a maturity of 5 years.\footnote{The duration of bonds in the model is $\frac{{(1+r)}^4}{{(1+r)}^4 - \delta}$} Finally, the Taylor rule coefficient on inflation is set to the standard value of $\phi_{\pi}=1.5$.

% Bibliography inclusion handled automatically by document structure
% Override \bibfilesfound to use only \texname.bib (contains all needed entries, avoids duplicates)
% \renewcommand{\bibfilesfound}{./\texname}
\smartbib

\pdfonly{
% execute this version if compiling with pdflatex
\end{document}
}
